---
title: "Spring 2023 data analysis, first trial"
Detail: 1:200 culture:LB, 24 hours
output: csv file
Author: Chadinthon (Minnie) Kittivorawong
Date: May 23, 2023
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, 
                      fig.height=3, fig.width=5, fig.align = "center")
library(tidyverse) 
library(infer)
library(broom)
library(lubridate)
library(readxl)
```

import data file
```{r}
abs_raw <- read_excel("2023-02-09 first trial full time.xlsx",
                          range = "B36:CU325")
colnames(abs_raw)[2] <- "temp"


lum_raw <- read_excel("2023-02-09 first trial full time.xlsx",
                          range = "B329:CU618")
colnames(lum_raw)[2] <- "temp"
```

Declare well
```{r}
MK67 = c("B3", "B4", "B5", "B6", "B7", "B8", "B9", "B10", "B11")
MK66 = c("C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10", "C11")
MK65 = c("D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11")
MK64 = c("E3", "E4", "E5", "E6", "E7", "E8", "E9", "E10", "E11")
MK63 = c("F3", "F4", "F5", "F6", "F7", "F8", "F9", "F10", "F11")
MK62 = c("G3", "G4", "G5", "G6", "G7", "G8", "G9", "G10", "G11")

# well that has different amount of LB/controls that get contaminated
contam = c("B2", "C2", "D2", "E2", "F2", "G2", "A5")

# well that have condensation at the start
conden = c("A6")

# well that contains no bacteria
control = c("A1", "A2", "A3", "A4", "A7", "A8", "A9","A10","A11", "A12","B1", "B12", "C1", "C12", "D1", "D12","E1", "E12", "F1", "F12", "G1", "G12", "H1","H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12")
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time <- abs_raw %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df <- abs_with_time %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time <- lum_raw %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df <- lum_with_time %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```


Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum <- cbind(abs_df, lum_df[4])

# assign strain/species/promoter group to each well
group_abs_lum <- abs_lum %>%
  mutate(# declare strain
         strain = case_when(
           well %in% MK62 ~ "MK62",
           well %in% MK63 ~ "MK63",
           well %in% MK64 ~ "MK64",
           well %in% MK65 ~ "MK65",
           well %in% MK66 ~ "MK66",
           well %in% MK67 ~ "MK67",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare species
         species = case_when(
           well %in% MK62 ~ "E. coli",
           well %in% MK63 ~ "E. coli",
           well %in% MK64 ~ "E. coli",
           well %in% MK65 ~ "E. clocae",
           well %in% MK66 ~ "E. clocae",
           well %in% MK67 ~ "E. clocae",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% MK62 ~ "no promoter",
           well %in% MK63 ~ "E. coli promoter",
           well %in% MK64 ~ "E. clocae promoter",
           well %in% MK65 ~ "no promoter",
           well %in% MK66 ~ "E. coli promoter",
           well %in% MK67 ~ "E. clocae promoter",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"))
```


normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean <- group_abs_lum %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi <- left_join(group_abs_lum, envi_mean, by = "Time")

# normalize abs and lum with background
normalize_abs_lum <- normalize_envi %>%
  mutate(absorbance = case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ 0), #assign number to 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ 0)) ##assign number to 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum <- normalize_abs_lum %>%
#normalized_lum <- normalize_envi %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data <- normalized_lum %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

Plot normalized luminescence out
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("normalized luminescence") +
    ggtitle("First trial") +
  facet_wrap(~strain)
```

plot normalized luminescence log scale
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("normalized luminescence (log10)") +
    ggtitle("First trial, log scale") +
    facet_wrap(~strain)
```

Plot raw abs out
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    xlab("Time (hours)") +
    ylab("raw absorbance") +
    ggtitle("Raw absorbance first trial") +
    facet_wrap(~strain)
```

Plot raw abs log scale
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    scale_y_continuous(trans = 'log10', limits = c(.01, 2)) +
    xlab("Time (min)") +
    ylab("raw absorbance (log10)") +
    ggtitle("Raw absorbance first trial") +
    facet_wrap(~strain)
```

Plot raw lum out
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence first trial") +
    facet_wrap(~strain)
```

Plot raw lum log scale
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  ggplot() + 
  geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
  scale_shape_manual(values=c(1, 2, 3))+
  scale_y_continuous(trans = 'log10') +
  xlab("Time (min)") +
  ylab("luminescence (log10)") +
  ggtitle("Raw luminescence first trial") +
  facet_wrap(~strain)
```

Plot lum at exponential growth
```{r}
correct_well_data %>%
  filter(strain != "control") %>%
  filter(absorbance > 0.05 & absorbance < 0.06) %>%
  ggplot(aes(x = strain, y = normalized, color = promoter)) +
  geom_point(aes(shape = species), position = "jitter", alpha = .4) +
  xlab("Strain") +
  ylab("luminescence (at abs = 0.05-0.0.06)") +
  ggtitle("Luminessence at exponential growth (First Trial)")
```

Calculate the lum at exponential growth phrase
```{r}
rep_lum <- correct_well_data %>%
  filter(strain != "control") %>%
  filter(absorbance > 0.05 & absorbance < 0.06) %>%
  group_by(strain) %>%
  summarise(repLum = mean(normalized),
            repLumsd = sd(normalized),
            repLumError = sd(normalized)/sqrt(n()),
            maxError = repLum + repLumError,
            minError = repLum - repLumError)

rep_lum <- rep_lum %>%
  mutate(# declare species
         species = case_when(
           strain == "MK62" ~ "E. coli",
           strain == "MK63" ~ "E. coli",
           strain == "MK64" ~ "E. coli",
           strain == "MK65" ~ "E. clocae",
           strain == "MK66" ~ "E. clocae",
           strain == "MK67" ~ "E. clocae"),
         # declare promoter
         promoter = case_when(
           strain == "MK62" ~ "no promoter",
           strain == "MK63" ~ "E. coli promoter",
           strain == "MK64" ~ "E. clocae promoter",
           strain == "MK65" ~ "no promoter",
           strain == "MK66" ~ "E. coli promoter",
           strain == "MK67" ~ "E. clocae promoter"),
         Experiment = "First trial")
```

Plot the lum at exponential growth phrase
```{r}
rep_lum %>%
  ggplot(aes(x = strain, y = repLum, color = promoter, shape = species)) +
  geom_point() +
  geom_errorbar(aes(ymin = repLum - repLumsd, ymax = repLum + repLumsd), width=.2,
                 position=position_dodge(0.05)) +
  scale_y_continuous(labels = function(y) format(y, scientific = TRUE)) + 
  xlab("Strain") +
  ylab("luminescence (at abs = 0.05-0.0.06)") +
  ggtitle("Luminessence at exponential growth (First trial)")
```

Export rep_lum as csv file
```{r}
write.csv(rep_lum, "/Users/student/Desktop/Harvey Mudd College/research spring 2023/First-trial-lum-expo.csv" ,row.names = FALSE)
```


