---
title: "Jo Thesis 37 Cdata analysis"
output: pdf_document
---

#all of the code for each trial individually is from minnie 
```{r global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, 
                      fig.height=3, fig.width=5, fig.align = "center")
library(tidyverse) 
library(infer)
library(broom)
library(lubridate)
library(readxl)
```

Declare well
```{r}
# saparate well by strain
JAA7 = c("B2", "B3", "B4", "B5", "B6", "B7", "B8", "B9", "B10","B11")
JAA13 = c("C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10","C11")
JAA17 = c("D2", "D3", "D4", "D5", "D6", "D7","D8", "D9", "D10", "D11")
JAA18 = c("E2", "E3", "E4", "E5", "E6", "E7", "E8", "E9", "E10", "E11")
JAA20 = c("F2", "F3", "F4", "F5", "F6", "F7", "F8", "F9", "F10" , "F11")
JAA21 = c("G2", "G3", "G4", "G5", "G6","G7", "G8", "G9", "G10", "G11")

# well that contains no bacteria but get contaminate
contam = c()

# well that have condensation at the start
conden = c()

# well that contains no bacteria
control = c("A1", "A2", "A3","A4", "A5", "A6", "A7", "A8", "A9", "A10","A11", "A12", "B1", "B12", "C1", "C12", "D1","D12", "E1", "E12", "F1", "F12","G1","G12","H1","H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11","H12" )
```

#37C Exp 1
import data file
```{r}
abs_raw_37_1 <- read_excel("2023-4-5 first 37.xlsx",
                          range = "B38:CU321")
colnames(abs_raw_37_1)[2] <- "temp"


lum_raw_37_1 <- read_excel("2023-4-5 first 37.xlsx",
                          range = "B907:CU1190")
colnames(lum_raw_37_1)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_37_1 <- abs_raw_37_1 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_37_1 <- abs_with_time_37_1 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_37_1 <- lum_raw_37_1 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_37_1 <- lum_with_time_37_1 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```


Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_37_1 <- cbind(abs_df_37_1, lum_df_37_1[4])

# assign strain/species/promoter group to each well
group_abs_lum_37_1 <- abs_lum_37_1 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_37_1 <- group_abs_lum_37_1 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_37_1 <- left_join(group_abs_lum_37_1, envi_mean_37_1, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_37_1 <- normalize_envi_37_1 %>%
  mutate(absorbance = 
           case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ NA), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ NA)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_37_1 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_37_1 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_37_1 <- normalize_abs_lum_37_1 %>%
#normalized_lum <- normalize_envi %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_37_1 <- normalized_lum_37_1 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```


#37C Exp 2
import data file
```{r}
abs_raw_37_2 <- read_excel("2023-4-7 second 37.xlsx",
                          range = "B40:CU316")
colnames(abs_raw_37_2)[2] <- "temp"


lum_raw_37_2 <- read_excel("2023-4-7 second 37.xlsx",
                          range = "B909:CU1185")
colnames(lum_raw_37_2)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_37_2 <- abs_raw_37_2 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_37_2 <- abs_with_time_37_2 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_37_2 <- lum_raw_37_2 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_37_2 <- lum_with_time_37_2 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```

```{r}
contam_37_2 = c("A11", "E1" )

# well that contains no bacteria
control_37_2 = c("A1", "A2", "A3","A4", "A5", "A6", "A7", "A8", "A9", "A10", "A12", "B1", "B12", "C1", "C12", "D1","D12", "E12", "F1", "F12","G1","G12","H1","H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11","H12" )
```

Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_37_2 <- cbind(abs_df_37_2, lum_df_37_2[4])

# assign strain/species/promoter group to each well
group_abs_lum_37_2 <- abs_lum_37_2 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam_37_2 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_37_2 ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam_37_2 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_37_2 ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam_37_2 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_37_2 ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_37_2 <- group_abs_lum_37_2 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_37_2 <- left_join(group_abs_lum_37_2, envi_mean_37_2, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_37_2 <- normalize_envi_37_2 %>%
  mutate(absorbance =
           case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ NA), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ NA)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_37_2 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_37_2 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_37_2 <- normalize_abs_lum_37_2 %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_37_2 <- normalized_lum_37_2 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```


#37C Exp 3
import data file
```{r}
abs_raw_37_3 <- read_excel("2023-4-10 third 37.xlsx",
                          range = "B40:CU307")
colnames(abs_raw_37_3)[2] <- "temp"


lum_raw_37_3 <- read_excel("2023-4-10 third 37.xlsx",
                          range = "B909:CU1176")
colnames(lum_raw_37_3)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_37_3 <- abs_raw_37_3 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_37_3 <- abs_with_time_37_3 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_37_3 <- lum_raw_37_3 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_37_3 <- lum_with_time_37_3 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```

which ones need to be taken out for this replicate
```{r}
contam_37_3 = c("H9","E1")

# well that contains no bacteria
control_37_3 = c("A1", "A2", "A3","A4", "A5", "A6", "A7", "A8", "A9", "A10", "A11", "A12", "B1", "B12", "C1", "C12", "D1","D12",  "E12", "F1", "F12","G1","G12","H1","H2", "H3", "H4", "H5", "H6", "H7", "H8",  "H10", "H11","H12" )
```


Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_37_3 <- cbind(abs_df_37_3, lum_df_37_3[4])

# assign strain/species/promoter group to each well
group_abs_lum_37_3 <- abs_lum_37_3 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam_37_3 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_37_3 ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam_37_3 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_37_3 ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam_37_3 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_37_3 ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_37_3 <- group_abs_lum_37_3 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_37_3 <- left_join(group_abs_lum_37_3, envi_mean_37_3, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_37_3 <- normalize_envi_37_3 %>%
  mutate(absorbance =  
           case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ NA), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ NA)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_37_3 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_37_3 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_37_3 <- normalize_abs_lum_37_3 %>%
#normalized_lum <- normalize_envi %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_37_3 <- normalized_lum_37_3 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

#New code - checking background & if growth looks normal for each well 
trial 1
```{r}
 data.frame(normalize_envi_37_1) %>%
  filter(strain == "control") %>% 
  ggplot(aes(x = Time, y = absorbance, colour = well)) +
  geom_point() +
  facet_wrap(~well)

 data.frame(normalize_envi_37_1) %>%
  filter(strain == "control") %>% 
  ggplot(aes(x = Time, y = absorbance, colour = well)) +
  geom_point() 
 
   data.frame(normalize_envi_37_1) %>% 
   filter(strain == "control") %>% 
    group_by(Time) %>%
   summarize(mean_back = mean(absorbance),
        Time = Time) %>% 
  ggplot(aes(x = Time, y = mean_back)) +
  geom_point() 
```
trial 2
```{r}
data.frame(normalize_envi_37_2) %>%
  filter(strain == "control") %>% 
  ggplot(aes(x = Time, y = absorbance, colour = well)) +
  geom_point() +
  facet_wrap(~well)

 data.frame(normalize_envi_37_2) %>%
  filter(strain == "control") %>% 
  ggplot(aes(x = Time, y = absorbance, colour = well)) +
  geom_point() 
 
   data.frame(normalize_envi_37_2) %>% 
   filter(strain == "control") %>% 
    group_by(Time) %>%
   summarize(mean_back = mean(absorbance),
        Time = Time) %>% 
  ggplot(aes(x = Time, y = mean_back)) +
  geom_point() 
```
trial 3
```{r}
data.frame(normalize_envi_37_3) %>%
  filter(strain == "control") %>% 
  ggplot(aes(x = Time, y = absorbance, colour = well)) +
  geom_point() +
  facet_wrap(~well)

 data.frame(normalize_envi_37_3) %>%
  filter(strain == "control") %>% 
  ggplot(aes(x = Time, y = absorbance, colour = well)) +
  geom_point() 
 
   data.frame(normalize_envi_37_3) %>% 
   filter(strain == "control") %>% 
    group_by(Time) %>%
   summarize(mean_back = mean(absorbance),
        Time = Time) %>% 
  ggplot(aes(x = Time, y = mean_back)) +
  geom_point() 
```


#Merging 37C Data into one df & find means
```{r}
correct_well_data_37_1 <- correct_well_data_37_1 %>% mutate(trial = "one")
correct_well_data_37_2 <- correct_well_data_37_2 %>% mutate(trial = "two")
correct_well_data_37_3 <- correct_well_data_37_3 %>% mutate(trial = "three")

correct_well_data_37 <- rbind(correct_well_data_37_1, correct_well_data_37_2)
correct_well_data_37 <- rbind(correct_well_data_37, correct_well_data_37_3)

correct_well_data_37_means <- correct_well_data_37 %>% group_by(strain, Time)%>%  
  summarise(mean_lum = mean(luminescence), 
            n = n(), 
            mean_abs = mean(absorbance)) %>% 
  dplyr::mutate(strain = case_when(
           strain == "JAA7" ~ "wt E.coli",
           strain == "JAA13" ~ "wt K.pneum",
           strain == "JAA17" ~ "E.c + E.c bolA",
           strain == "JAA18" ~ "E.c + K.p bolA",
           strain == "JAA20" ~ "K.p + E.c bolA",
           strain == "JAA21" ~ "K.p + K.p bolA",
           strain == "control" ~ "control"))
```
#plot lum & abs (just normalized to background) for each strain over time
```{r}
stat_lum_plot <- data.frame(correct_well_data_37_means) %>%
  filter(strain != "control") %>%
  dplyr::mutate(strain = fct_relevel(
    strain,
    "wt E.coli",
    "wt K.pneum",
    "E.c + E.c bolA",
    "E.c + K.p bolA",
    "K.p + E.c bolA",
    "K.p + K.p bolA")) %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = mean_lum, colour = strain)) +
    xlab("Time (hours)") +
    ylab("Luminescence") +
    ggtitle("Mean Normalized Luminescence at 37C") +
    facet_grid(~strain)
stat_lum_plot
ggsave("37C lum plot.png", stat_lum_plot, height = 4, width = 9)
```

```{r}
stat_abs_plot <- data.frame(correct_well_data_37_means) %>%
  filter(strain != "control") %>%
  dplyr::mutate(strain = fct_relevel(
    strain,
    "wt E.coli",
    "wt K.pneum",
    "E.c + E.c bolA",
    "E.c + K.p bolA",
    "K.p + E.c bolA",
    "K.p + K.p bolA")) %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = mean_abs, colour = strain)) +
    xlab("Time (hours)") +
    ylab("Log Transformed Absorbance") +
    ggtitle("Mean Normalized absorbance 37C") +
    facet_grid(~strain) +
  scale_y_log10() +
  NULL
stat_abs_plot
#ggsave("37C abs plot.png", stat_abs_plot, height = 4, width = 9)
```

#Making regression to approx doubling time 
picking exponential times 
```{r}
reg_times_37 <- correct_well_data_37 %>% 
  filter(Time > 1 &
           Time < 5) %>%
  mutate(log_abs = log2(absorbance))
```

regressions on all strains individually
```{r}
jaa7_37 <- data.frame(reg_times_37) %>% filter(strain == "JAA7", 
                                               !is.na(absorbance)) %>%
ggplot(jaa7_37, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
lm(log_abs ~ Time, data = jaa7_37)

jaa13_37 <- data.frame(reg_times_37) %>% filter(strain == "JAA13")
ggplot(jaa13_37, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
lm(log_abs ~ Time, data = jaa13_37)

jaa17_37 <- data.frame(reg_times_37) %>% filter(strain == "JAA17")
ggplot(jaa17_37, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
lm(log_abs ~ Time, data = jaa17_37)

jaa18_37 <- data.frame(reg_times_37) %>% filter(strain == "JAA18")
ggplot(jaa18_37, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm) #this one doesn't look at good
lm(log_abs ~ Time, data = jaa18_37)

jaa20_37 <- data.frame(reg_times_37) %>% filter(strain == "JAA20")
ggplot(jaa20_37, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
lm(log_abs ~ Time, data = jaa20_37)

jaa21_37 <- data.frame(reg_times_37) %>% filter(strain == "JAA21")
ggplot(jaa21_37, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
lm(log_abs ~ Time, data = jaa21_37)
```
```{r}
 new_Time <- data.frame(Time = seq(from = 1.083333, to = 15, by = 0.083334))
jaa7_37_est <- new_Time %>%
  mutate(N_t = -9.865 + (1/1.804)*Time)
```


#picking od & normalizing
3 hr exp to avoid any weirdness at the beginning of the experiment 
```{r}
exp_37 <- correct_well_data_37 %>% 
  filter(Time > 2.9 &
           Time < 3.1) %>%
  mutate(group = "37C exp")
```
10 hr stat
```{r}
stat_37 <- correct_well_data_37 %>% 
  filter(Time > 14.9 &
           Time < 15.1) %>%
  
  mutate(group = "37C stat")
```
```{r}
all_37 <- rbind(exp_37,stat_37)
write.csv(all_37, "stat & exp 37c.csv" )
```
```{r}
correct_well_data_37 %>% filter(Time > .9 &
           Time < 1.1)  %>% filter(strain != "control")
```


#make tables for stat_exp_plot_code to compare ods across growth conditions

```{r}
exp_37 %>% group_by(Time, strain) %>% filter(!is.na(luminescence), !is.na(absorbance), strain != "control", strain != "JAA7", strain != "JAA13"    )%>%
  summarise(mean_lum = mean(luminescence), 
            mean_abs = mean(absorbance), 
            mean_norm = mean(normalized))


stat_37 %>% group_by(Time, strain) %>% filter(!is.na(luminescence), !is.na(absorbance), strain != "control", strain != "JAA7", strain != "JAA13"    )%>%
  summarise(mean_lum = mean(luminescence), 
            mean_abs = mean(absorbance), 
            mean_norm = mean(normalized))
```

