---
title: "Jo Thesis 15C Data analysis"
author: "Josephine Adams"
date: "2023-04-10"
output: html_document
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, 
                      fig.height=3, fig.width=5, fig.align = "center")
library(tidyverse) 
library(infer)
library(broom)
library(lubridate)
library(readxl)
```

Declare well
```{r}
# separate well by strain
JAA7 = c("B2", "B3", "B4", "B5", "B6", "B7", "B8", "B9", "B10","B11")
JAA13 = c("C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10","C11")
JAA17 = c("D2", "D3", "D4", "D5", "D6", "D7","D8", "D9", "D10", "D11")
JAA18 = c("E2", "E3", "E4", "E5", "E6", "E7", "E8", "E9", "E10", "E11")
JAA20 = c("F2", "F3", "F4", "F5", "F6", "F7", "F8", "F9", "F10" , "F11")
JAA21 = c("G2", "G3", "G4", "G5", "G6","G7", "G8", "G9", "G10", "G11")

# well that contains no bacteria but get contaminate
contam = c()

# well that have condensation at the start
conden = c()

# well that contains no bacteria
control = c("A1", "A2", "A3","A4", "A5", "A6", "A7", "A8", "A9", "A10","A11", "A12", "B1", "B12", "C1", "C12", "D1","D12", "E1", "E12", "F1", "F12","G1","G12","H1","H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11","H12" )
```

#15C Exp 1
import data file
```{r}
abs_raw_15_1 <- read_excel("2023-4-6 first 15.xlsx",
                          range = "B40:CU332")
colnames(abs_raw_15_1)[2] <- "temp"


lum_raw_15_1 <- read_excel("2023-4-6 first 15.xlsx",
                          range = "B909:CU1201")
colnames(lum_raw_15_1)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_15_1 <- abs_raw_15_1 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_15_1 <- abs_with_time_15_1 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_15_1 <- lum_raw_15_1 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_15_1 <- lum_with_time_15_1 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```

which ones to remove for this samle
```{r}
JAA13_15_1 = c("C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10")
JAA21_15_1 = c("G2", "G3", "G4", "G5", "G6","G7", "G8", "G9", "G10")

# well that contains no bacteria but get contaminate
contam_15_1 = c("C11", "G11" )
```

Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_15_1 <- cbind(abs_df_15_1, lum_df_15_1[4])

# assign strain/species/promoter group to each well
group_abs_lum_15_1 <- abs_lum_15_1 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13_15_1 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21_15_1 ~ "JAA21",
           well %in% contam_15_1 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13_15_1 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21_15_1 ~ "K. pneumoniae",
           well %in% contam_15_1 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13_15_1 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21_15_1 ~ "K.pneum promoter",
           well %in% contam_15_1 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_15_1 <- group_abs_lum_15_1 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_15_1 <- left_join(group_abs_lum_15_1, envi_mean_15_1, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_15_1 <- normalize_envi_15_1 %>%
  mutate(absorbance = case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ 0.0000001), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ 0.0000001)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_15_1 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_15_1 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_15_1 <- normalize_abs_lum_15_1 %>%
#normalized_lum <- normalize_envi %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_15_1 <- normalized_lum_15_1 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

Plot normalized luminescence out
```{r}
correct_well_data_15_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("normalized luminescence") +
    ggtitle("15C Exp 1") +
  facet_wrap(~strain)
```

plot normalized luminescence log scale
```{r}
correct_well_data_15_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("normalized luminescence (log10)") +
    ggtitle("Third trial, log scale") +
    facet_wrap(~strain)
```

Plot raw abs out
```{r}
correct_well_data_15_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (hours)") +
    ylab("raw absorbance") +
    ggtitle("Raw absorbance third trial") +
    facet_wrap(~strain)
```
JAA18 basically didn't grow. Should I remove it. I think it may not ever grow.  
 
Plot raw abs log scale
```{r}
correct_well_data_15_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("raw absorbance (log10)") +
    ggtitle("Raw absorbance third trial") +
    facet_wrap(~strain)
```

Plot raw lum out
```{r}
correct_well_data_15_1 %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence third trial") +
    facet_wrap(~strain)
```

Plot raw lum log scale
```{r}
correct_well_data_15_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
  geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
  scale_shape_manual(values=c(1, 2, 3))+
  geom_point(mapping = aes(x = Time, y = mean_lum)) +
  scale_y_continuous(trans = 'log10') +
  xlab("Time (min)") +
  ylab("luminescence (log10)") +
  ggtitle("Raw luminescence third trial") +
  facet_wrap(~strain)
```

#15C Exp 2
import data file
```{r}
abs_raw_15_2 <- read_excel("2023-4-8 second 15.xlsx",
                          range = "B42:CU330")
colnames(abs_raw_15_2)[2] <- "temp"


lum_raw_15_2 <- read_excel("2023-4-8 second 15.xlsx",
                          range = "B911:CU1199")
colnames(lum_raw_15_2)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_15_2 <- abs_raw_15_2 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_15_2 <- abs_with_time_15_2 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_15_2 <- lum_raw_15_2 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_15_2 <- lum_with_time_15_2 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```

what to remove for this replicate 
```{r}
JAA20_15_2 = c("F2", "F3", "F4", "F5", "F6", "F7", "F8", "F9", "F10" )

# well that contains no bacteria but get contaminate
contam_15_2 = c("F11")
```

Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_15_2 <- cbind(abs_df_15_2, lum_df_15_2[4])

# assign strain/species/promoter group to each well
group_abs_lum_15_2 <- abs_lum_15_2 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20_15_2 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam_15_2 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20_15_2 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam_15_2 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20_15_2 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam_15_2 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_15_2 <- group_abs_lum_15_2 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_15_2 <- left_join(group_abs_lum_15_2, envi_mean_15_2, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_15_2 <- normalize_envi_15_2 %>%
  mutate(absorbance = case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ 0.0000001), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ 0.0000001)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_15_2 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_15_2 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_15_2 <- normalize_abs_lum_15_2 %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_15_2 <- normalized_lum_15_2 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

Plot normalized luminescence out
```{r}
correct_well_data_15_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("normalized luminescence") +
    ggtitle("15C Exp 2") +
  facet_wrap(~strain)
```

plot normalized luminescence log scale
```{r}
correct_well_data_15_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("normalized luminescence (log10)") +
    ggtitle("15 Exp 2, log scale") +
    facet_wrap(~strain)
```

Plot raw abs out
```{r}
correct_well_data_15_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (hours)") +
    ylab("raw absorbance") +
    ggtitle("Raw absorbance 15 Exp 2") +
    facet_wrap(~strain)
```

Plot raw abs log scale
```{r}
correct_well_data_15_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("raw absorbance (log10)") +
    ggtitle("Raw absorbance 15 Exp 2") +
    facet_wrap(~strain)
```

Plot raw lum out
```{r}
correct_well_data_15_2 %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence 15 Exp 2") +
    facet_wrap(~strain)
```

Plot raw lum log scale
```{r}
correct_well_data_15_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
  geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
  scale_shape_manual(values=c(1, 2, 3))+
  geom_point(mapping = aes(x = Time, y = mean_lum)) +
  scale_y_continuous(trans = 'log10') +
  xlab("Time (min)") +
  ylab("luminescence (log10)") +
  ggtitle("Raw luminescence log 15 Exp 2 ") +
  facet_wrap(~strain)
```

#15C Exp 3
import data file
```{r}
abs_raw_15_3 <- read_excel("2023-4-9 third 14.xlsx",
                          range = "B42:CU319")
colnames(abs_raw_15_3)[2] <- "temp"


lum_raw_15_3 <- read_excel("2023-4-9 third 14.xlsx",
                          range = "B911:CU1188")
colnames(lum_raw_15_3)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_15_3 <- abs_raw_15_3 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_15_3 <- abs_with_time_15_3 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_15_3 <- lum_raw_15_3 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_15_3 <- lum_with_time_15_3 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```

```{r}
JAA13_15_3 = c( "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10","C11")
JAA17_15_3 = c("D2", "D3", "D4", "D5", "D6", "D7","D8", "D9", "D10")

# well that contains no bacteria but get contaminate
contam_15_3 = c("A2", "A3","C2","D11")
control_15_3 = c("A1", "A4", "A5", "A6", "A7", "A8", "A9", "A10","A11", "A12", "B1", "B12", "C1", "C12", "D1","D12", "E1", "E12", "F1", "F12","G1","G12","H1","H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11","H12" )
```


Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_15_3 <- cbind(abs_df_15_3, lum_df_15_3[4])

# assign strain/species/promoter group to each well
group_abs_lum_15_3 <- abs_lum_15_3 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13_15_3 ~ "JAA13",
           well %in% JAA17_15_3 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam_15_3 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_15_3 ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13_15_3 ~ "K. pneumoniae",
           well %in% JAA17_15_3 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam_15_3 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_15_3 ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13_15_3 ~ "no promoter",
           well %in% JAA17_15_3 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam_15_3 ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control_15_3 ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_15_3 <- group_abs_lum_15_3 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_15_3 <- left_join(group_abs_lum_15_3, envi_mean_15_3, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_15_3 <- normalize_envi_15_3 %>%
  mutate(absorbance = case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ 0.0000001), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ 0.0000001)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_15_3 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_15_3 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_15_3 <- normalize_abs_lum_15_3 %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_15_3 <- normalized_lum_15_3 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

Plot normalized luminescence out
```{r}
correct_well_data_15_3 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("normalized luminescence") +
    ggtitle("15C Exp 2") +
  facet_wrap(~strain)
```

plot normalized luminescence log scale
```{r}
correct_well_data_15_3 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("normalized luminescence (log10)") +
    ggtitle("15 Exp 2, log scale") +
    facet_wrap(~strain)
```

Plot raw abs out
```{r}
correct_well_data_15_3 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (hours)") +
    ylab("raw absorbance") +
    ggtitle("Raw absorbance 15 Exp 2") +
    facet_wrap(~strain)
```

Plot raw abs log scale
```{r}
correct_well_data_15_3 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("raw absorbance (log10)") +
    ggtitle("Raw absorbance 15 Exp 2") +
    facet_wrap(~strain)
```

Plot raw lum out
```{r}
correct_well_data_15_3 %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence 15 Exp 2") +
    facet_wrap(~strain)
```

Plot raw lum log scale
```{r}
correct_well_data_15_3 %>%
  filter(strain != "control") %>%
  ggplot() + 
  geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
  scale_shape_manual(values=c(1, 2, 3))+
  geom_point(mapping = aes(x = Time, y = mean_lum)) +
  scale_y_continuous(trans = 'log10') +
  xlab("Time (min)") +
  ylab("luminescence (log10)") +
  ggtitle("Raw luminescence log 15 Exp 2 ") +
  facet_wrap(~strain)
```
#Merging 15C Data 
```{r}
correct_well_data_15 <- rbind(correct_well_data_15_1, correct_well_data_15_2)
correct_well_data_15 <- rbind(correct_well_data_15, correct_well_data_15_3)

correct_well_data_15_means <- correct_well_data_15 %>% group_by(strain, Time)%>%  
  summarise(mean_lum = mean(luminescence), 
            n = n(), 
            mean_abs = mean(absorbance)) %>% 
  dplyr::mutate(strain = case_when(
           strain == "JAA7" ~ "wt E.coli",
           strain == "JAA13" ~ "wt K.pneum",
           strain == "JAA17" ~ "E.c + E.c bolA",
           strain == "JAA18" ~ "E.c + K.p bolA",
           strain == "JAA20" ~ "K.p + E.c bolA",
           strain == "JAA21" ~ "K.p + K.p bolA",
           strain == "control" ~ "control"))
```

```{r}
cold_lum_plot <- data.frame(correct_well_data_15_means) %>%
  filter(strain != "control") %>%
  dplyr::mutate(strain = fct_relevel(
    strain,
    "wt E.coli",
    "wt K.pneum",
    "E.c + E.c bolA",
    "E.c + K.p bolA",
    "K.p + E.c bolA",
    "K.p + K.p bolA")) %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = mean_lum, colour = strain)) +
    xlab("Time (hours)") +
    ylab("Luminescence") +
    ggtitle("Mean luminescence 15C") +
    facet_grid(~strain)
cold_lum_plot
ggsave("15C lum plot.png", cold_lum_plot, height = 4, width = 9)
```

```{r}
cold_abs_plot <- data.frame(correct_well_data_15_means) %>%
  filter(strain != "control") %>%
  dplyr::mutate(strain = fct_relevel(
    strain,
    "wt E.coli",
    "wt K.pneum",
    "E.c + E.c bolA",
    "E.c + K.p bolA",
    "K.p + E.c bolA",
    "K.p + K.p bolA")) %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = mean_abs, colour = strain)) +
    xlab("Time (hours)") +
    ylab("Log Transformed Absorbance") +
    ggtitle("Mean absorbance 15C") +
    facet_grid(~strain) +
  scale_y_log10()
cold_abs_plot
ggsave("15C abs plot.png", cold_abs_plot, height = 4, width = 9)
```

#Making regression to approx doubling time 
```{r}
reg_times_15 <- correct_well_data_15 %>% 
  filter(Time > 3 &
           Time < 15) %>%
  mutate(log_abs = log2(absorbance))
```
regressions on all strains individually
```{r}
jaa7_15 <- data.frame(reg_times_15) %>% filter(strain == "JAA7")
ggplot(jaa7_15, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
jaa7_model_15 <- lm(log_abs ~ Time, data = jaa7_15)

jaa13_15 <- data.frame(reg_times_15) %>% filter(strain == "JAA13")
ggplot(jaa13_15, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
jaa13_model_15 <- lm(log_abs ~ Time, data = jaa13_15)

jaa17_15 <- data.frame(reg_times_15) %>% filter(strain == "JAA17")
ggplot(jaa17_15, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
jaa17_model_15 <- lm(log_abs ~ Time, data = jaa17_15)

jaa18_15 <- data.frame(reg_times_15) %>% filter(strain == "JAA18")
ggplot(jaa18_15, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm) #this one doesn't look at good
jaa18_model_15 <- lm(log_abs ~ Time, data = jaa18_15)

jaa20_15 <- data.frame(reg_times_15) %>% filter(strain == "JAA20")
ggplot(jaa20_15, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
jaa20_model_15 <- lm(log_abs ~ Time, data = jaa20_15)

jaa21_15 <- data.frame(reg_times_15) %>% filter(strain == "JAA21")
ggplot(jaa21_15, aes(x =Time, y=log_abs)) +
  geom_point() +
  stat_smooth(method = lm)
jaa21_model_15 <- lm(log_abs ~ Time, data = jaa21_15)
```
make a new table with doubling time estimations
```{r}
jaa7_15_est <- jaa7_15 %>%
  mutate(N_t = -9.223 + (1/0.3397)*Time)

jaa13_15_est <- jaa13_15 %>%
  mutate(N_t = -8.4561 + (1/0.4116)*Time)

doub_est <- rbind(jaa7_15_est, jaa13_15_est)

jaa17_15_est <- jaa17_15 %>%
  mutate(N_t = -9.489 + (1/0.330)*Time)

doub_est <- rbind(doub_est, jaa17_15_est)

jaa18_15_est <- jaa18_15 %>%
  mutate(N_t = -16.8586  + (1/0.4537)*Time)

doub_est <- rbind(doub_est, jaa18_15_est)

jaa20_15_est <- jaa20_15 %>%
  mutate(N_t = -7.6679 + (1/0.3719)*Time)

doub_est <- rbind(doub_est, jaa20_15_est)

jaa21_15_est <- jaa21_15 %>%
  mutate(N_t = -7.9694  + (1/0.4268)*Time)

doub_est <- rbind(doub_est, jaa21_15_est)
```
plot & get means
```{r}
doub_est %>%
  ggplot()+
  geom_point(aes(x = Time, y = 2^(N_t)))+
  facet_wrap(~strain, scales = "free_y")+
  #facet_grid(~strain)+
  ylab("Estimated amount of bacteria")
  
```



#picking od 
3 hr exp to let it recover from
```{r}
exp_15 <- correct_well_data_15 %>% 
  filter(Time > 2.9 &
           Time < 3.1) %>%
  mutate(group = "15 exp")
write.csv(exp_15, "exponential 15c.csv" )
```

