---
title: "Jo Thesis data analysis"
output: pdf_document
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, 
                      fig.height=3, fig.width=5, fig.align = "center")
library(tidyverse) 
library(infer)
library(broom)
library(lubridate)
library(readxl)
```

Declare well
```{r}
# saparate well by strain
JAA7 = c("B2", "B3", "B4", "B5", "B6", "B7", "B8", "B9", "B10","B11")
JAA13 = c("C2", "C3", "C4", "C5", "C6", "C7", "C8", "C9", "C10","C11")
JAA17 = c("D2", "D3", "D4", "D5", "D6", "D7","D8", "D9", "D10", "D11")
JAA18 = c("E2", "E3", "E4", "E5", "E6", "E7", "E8", "E9", "E10", "E11")
JAA20 = c("F2", "F3", "F4", "F5", "F6", "F7", "F8", "F9", "F10" , "F11")
JAA21 = c("G2", "G3", "G4", "G5", "G6","G7", "G8", "G9", "G10", "G11")

# well that contains no bacteria but get contaminate
contam = c()

# well that have condensation at the start
conden = c()

# well that contains no bacteria
control = c("A1", "A2", "A3","A4", "A5", "A6", "A7", "A8", "A9", "A10","A11", "A12", "B1", "B12", "C1", "C12", "D1","D12", "E1", "E12", "F1", "F12","G1","G12","H1","H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11","H12" )
```

#37C Exp 1
import data file
```{r}
abs_raw_37_1 <- read_excel("2023-4-5 first 37.xlsx",
                          range = "B38:CU321")
colnames(abs_raw_37_1)[2] <- "temp"


lum_raw_37_1 <- read_excel("2023-4-5 first 37.xlsx",
                          range = "B907:CU1190")
colnames(lum_raw_37_1)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_37_1 <- abs_raw_37_1 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_37_1 <- abs_with_time_37_1 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_37_1 <- lum_raw_37_1 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_37_1 <- lum_with_time_37_1 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```


Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_37_1 <- cbind(abs_df_37_1, lum_df_37_1[4])

# assign strain/species/promoter group to each well
group_abs_lum_37_1 <- abs_lum_37_1 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_37_1 <- group_abs_lum_37_1 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_37_1 <- left_join(group_abs_lum_37_1, envi_mean_37_1, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_37_1 <- normalize_envi_37_1 %>%
  mutate(absorbance = case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ 0.0000001), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ 0.0000001)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_37_1 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_37_1 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_37_1 <- normalize_abs_lum_37_1 %>%
#normalized_lum <- normalize_envi %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_37_1 <- normalized_lum_37_1 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

Plot normalized luminescence out
```{r}
correct_well_data_37_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("normalized luminescence") +
    ggtitle("37C Exp 1") +
  facet_wrap(~strain)
```

plot normalized luminescence log scale
```{r}
correct_well_data_37_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("normalized luminescence (log10)") +
    ggtitle("Third trial, log scale") +
    facet_wrap(~strain)
```

Plot raw abs out
```{r}
correct_well_data_37_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (hours)") +
    ylab("raw absorbance") +
    ggtitle("Raw absorbance third trial") +
    facet_wrap(~strain)
```

Plot raw abs log scale
```{r}
correct_well_data_37_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("raw absorbance (log10)") +
    ggtitle("Raw absorbance third trial") +
    facet_wrap(~strain)
```

Plot raw lum out
```{r}
correct_well_data_37_1 %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence third trial") +
    facet_wrap(~strain)
```

Plot raw lum log scale
```{r}
correct_well_data_37_1 %>%
  filter(strain != "control") %>%
  ggplot() + 
  geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
  scale_shape_manual(values=c(1, 2, 3))+
  geom_point(mapping = aes(x = Time, y = mean_lum)) +
  scale_y_continuous(trans = 'log10') +
  xlab("Time (min)") +
  ylab("luminescence (log10)") +
  ggtitle("Raw luminescence third trial") +
  facet_wrap(~strain)
```
#37C Exp 2
import data file
```{r}
abs_raw_37_2 <- read_excel("2023-4-7 second 37.xlsx",
                          range = "B40:CU316")
colnames(abs_raw_37_2)[2] <- "temp"


lum_raw_37_2 <- read_excel("2023-4-5 first 37.xlsx",
                          range = "B907:CU1183")
colnames(lum_raw_37_2)[2] <- "temp"
```

normalized time/pivot table
```{r}
# normalized time into minutes
abs_with_time_37_2 <- abs_raw_37_2 %>%
  mutate(Time = Time[1] %--% Time, # use time interval to normalize time
         Time = as.numeric(Time, unit = "hours") )

# pivot absorbance table
abs_df_37_2 <- abs_with_time_37_2 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "absorbance"
  )

# normalized time into minutes
lum_with_time_37_2 <- lum_raw_37_2 %>%
  mutate(Time = Time[1] %--% Time,
         Time = as.numeric(Time, unit = "hours"))

# pivot luminescence table
lum_df_37_2 <- lum_with_time_37_2 %>% 
  pivot_longer(
    cols = !Time & !temp, 
    names_to = "well", 
    values_to = "luminescence"
  )
```


Combine abs and lum table and declare strain/species/promoter group
```{r}
# combine abs and lum dataframe
abs_lum_37_2 <- cbind(abs_df_37_2, lum_df_37_2[4])

# assign strain/species/promoter group to each well
group_abs_lum_37_2 <- abs_lum_37_2 %>%
  mutate(# declare strain
         strain = case_when(
           well %in% JAA7 ~ "JAA7",
           well %in% JAA13 ~ "JAA13",
           well %in% JAA17 ~ "JAA17",
           well %in% JAA18 ~ "JAA18",
           well %in% JAA20 ~ "JAA20",
           well %in% JAA21 ~ "JAA21",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare species
         species = case_when(
           well %in% JAA7 ~ "E. coli",
           well %in% JAA13 ~ "K. pneumoniae",
           well %in% JAA17 ~ "E. coli",
           well %in% JAA18 ~ "E. coli",
           well %in% JAA20 ~ "K. pneumoniae",
           well %in% JAA21 ~ "K. pneumoniae",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"),
         # declare promoter
         promoter = case_when(
           well %in% JAA7 ~ "no promoter",
           well %in% JAA13 ~ "no promoter",
           well %in% JAA17 ~ "E. coli promoter",
           well %in% JAA18 ~ "K.pneum promoter",
           well %in% JAA20 ~ "E. coli promoter",
           well %in% JAA21 ~ "K.pneum promoter",
           well %in% contam ~ "contaminate",
           well %in% conden ~ "condensated",
           well %in% control ~ "control"))
```

normalized absorbance and luminescence with backgroud
```{r}
# get background abs and lum at each time
envi_mean_37_2 <- group_abs_lum_37_2 %>%
  filter(strain == "control") %>%
  group_by(Time) %>%
  summarize(mean_abs = mean(absorbance),
            mean_lum = mean(luminescence))

# combine background abs and lum data with experiment data
normalize_envi_37_2 <- left_join(group_abs_lum_37_2, envi_mean_37_2, by = "Time")

# normalize abs and lum with background
normalize_abs_lum_37_2 <- normalize_envi_37_2 %>%
  mutate(absorbance = case_when(
           absorbance > mean_abs ~ absorbance - mean_abs,
           absorbance <= mean_abs ~ 0.0000001), #assign number to prevent 0
         luminescence = case_when(
           luminescence > mean_lum ~ luminescence - mean_lum,
           luminescence <= mean_lum ~ 0.0000001)) ##assign number to prevent 0
```

Plot envi_mean out
```{r}
# plot background abs
envi_mean_37_2 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (min)") +
    ylab("background abs") +
    ggtitle("envi_mean, abs")

# plot background lum
envi_mean_37_2 %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (min)") +
    ylab("background lum") +
    ggtitle("envi_mean, lum")
```

normalized luminescence with absorbance
```{r}
# normalised the luminescence to the absorbance
normalized_lum_37_2 <- normalize_abs_lum_37_2 %>%
  mutate(normalized = luminescence/absorbance)
```

Choose only used well, don't use the well that has contamination potential or have condensation at the start
```{r}
correct_well_data_37_2 <- normalized_lum_37_2 %>%
  filter(strain != "contaminate") %>%
  filter(strain != "condensated")
```

Plot normalized luminescence out
```{r}
correct_well_data_37_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    xlab("Time (hours)") +
    ylab("normalized luminescence") +
    ggtitle("37C Exp 2") +
  facet_wrap(~strain)
```

plot normalized luminescence log scale
```{r}
correct_well_data_37_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = normalized, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("normalized luminescence (log10)") +
    ggtitle("37 Exp 2, log scale") +
    facet_wrap(~strain)
```

Plot raw abs out
```{r}
correct_well_data_37_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    xlab("Time (hours)") +
    ylab("raw absorbance") +
    ggtitle("Raw absorbance 37 Exp 2") +
    facet_wrap(~strain)
```

Plot raw abs log scale
```{r}
correct_well_data_37_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
    geom_point(mapping = aes(x = Time, y = absorbance, color = strain, shape = promoter))+
    scale_shape_manual(values=c(1, 2, 3, 4))+
    geom_point(mapping = aes(x = Time, y = mean_abs)) +
    scale_y_continuous(trans = 'log10') +
    xlab("Time (min)") +
    ylab("raw absorbance (log10)") +
    ggtitle("Raw absorbance 37 Exp 2") +
    facet_wrap(~strain)
```

Plot raw lum out
```{r}
correct_well_data_37_2 %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
    scale_shape_manual(values=c(1, 2, 3))+
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence 37 Exp 2") +
    facet_wrap(~strain)
```

Plot raw lum log scale
```{r}
correct_well_data_37_2 %>%
  filter(strain != "control") %>%
  ggplot() + 
  geom_point(mapping = aes(x = Time, y = luminescence, shape = species, color = strain)) +
  scale_shape_manual(values=c(1, 2, 3))+
  geom_point(mapping = aes(x = Time, y = mean_lum)) +
  scale_y_continuous(trans = 'log10') +
  xlab("Time (min)") +
  ylab("luminescence (log10)") +
  ggtitle("Raw luminescence log 37 Exp 2 ") +
  facet_wrap(~strain)
```
#Merging 37C Data 
```{r}
correct_well_data_37 <- rbind(correct_well_data_37_1, correct_well_data_37_2)

correct_well_data_37_means <- correct_well_data_37 %>% group_by(strain, Time)%>%  
  summarise(mean_lum = mean(luminescence))
```

```{r}
correct_well_data_37_means %>%
  filter(strain != "control") %>%
  ggplot() +  
    geom_point(mapping = aes(x = Time, y = mean_lum)) +
    xlab("Time (hours)") +
    ylab("luminescence") +
    ggtitle("Raw luminescence 37C") +
    facet_wrap(~strain)
```

